---
title: "09-new-snp-allele-freq"
author: "Aileen"
date: "2025-05-07"
output: html_document
---

---

## Load Libraries

```{r}
library(tidyverse)
library(ggplot2)
library(readr)
library(patchwork)
```

---

## Load Data (Header trimmed via `grep -v "##"` already)

```{r}
Sarisha_snps_only <- read_tsv("data/new_filtered_vcfs/sarisha_filtered_clean.vcf")
early_snps <- read_tsv("data/new_filtered_vcfs/early_pool_filtered_clean.vcf")
late_snps <- read_tsv("data/new_filtered_vcfs/late_pool_filtered_clean.vcf")

```

---

## Extract Good SNPs from Sarisha

```{r}
sarisha <- Sarisha_snps_only %>%
  rename(CHR = `#CHROM`) %>%
  select(CHR, POS, unknown) %>%
  separate(unknown, into = c("GT", NA, "AD", NA, NA, NA, NA), sep = ":", remove = FALSE) %>%
  separate(AD, into = c("AD_1", "AD_2"), sep = ",", convert = TRUE)

good_snps <- sarisha %>%
  filter(GT == "1/1", AD_1 == "0", AD_2 > 20) %>%
  mutate(CHR_POS = paste(CHR, POS, sep = "_")) %>%
  select(CHR_POS)

#write_tsv(good_snps, "output_tables/sarisha_good_snps.tsv")
```

---

## Filter Pool Datasets

```{r}
early_snps_filtered <- early_snps %>%
  rename(CHR = `#CHROM`) %>%
  mutate(CHR_POS = paste(CHR, POS, sep = "_")) %>%
  semi_join(good_snps, by = "CHR_POS")

late_snps_filtered <- late_snps %>%
  rename(CHR = `#CHROM`) %>%
  mutate(CHR_POS = paste(CHR, POS, sep = "_")) %>%
  semi_join(good_snps, by = "CHR_POS")
```

---

## Calculate Allele Frequencies
Calculate AO (alternate allele observations) divided by DP (total observation counts) to get allele frequency

```{r}
calc_allele_freq <- function(df) {
  new_df <- df %>%
    select(CHR_POS, unknown) %>%
    separate(unknown, into = c(NA, "DP", NA, NA, NA, "AO", NA, NA), sep = ":", convert = TRUE, fill = "right") %>%
    dplyr::mutate(AO = as.numeric(AO),
                DP = as.numeric(DP),
                allele_freq = (AO / DP)
                )
  return(new_df)
}

early_snp_allelefreq <- calc_allele_freq(early_snps_filtered)
late_snp_allelefreq <- calc_allele_freq(late_snps_filtered)

```

---

##  Combine for Plotting

```{r}
rapa_chr <- paste0("A0", 1:9) %>% append("A10")

format_for_plot <- function(df, pool_label) {
  df %>%
    separate(CHR_POS, into = c("CHR", "POS"), sep = "_") %>%
    filter(CHR %in% rapa_chr) %>%
    mutate(pool = pool_label, POSn = as.numeric(POS))
}

plot_early <- format_for_plot(early_snp_allelefreq, "early")
plot_late <- format_for_plot(late_snp_allelefreq, "late")

plot_all <- bind_rows(plot_early, plot_late)
```

---

## Bin SNPs and Calculate Density

```{r}
breaks <- seq(1, 60281114, 50000)
labels <- seq(50000, 60281114, 50000)

plot_all_binned <- plot_all %>%
  mutate(bin = cut(POSn, breaks = breaks, labels = labels))

plot_all_SNP_density <- plot_all_binned %>%
  group_by(bin, CHR, pool) %>%
  summarise(SNP_density = mean(allele_freq, na.rm = TRUE), .groups = "drop") %>%
  filter(!is.na(SNP_density))
```

---

## Save Binned Data

```{r}
write_csv(plot_all_binned, "output_tables/plot_all_binned.csv")
write_csv(plot_all_SNP_density, "output_tables/SNP_density.csv")

#plot_all_binned <- read_csv("output_tables/plot_all_binned.csv")
#plot_all_SNP_density <- read_csv("output_tables/SNP_density.csv")
```

---

## Plot SNP Density by Pool and Chromosome

```{r fig.height=20, fig.width=12}
SNP_density_plot <- plot_all_SNP_density %>%
  ggplot(aes(x = bin, y = SNP_density, color = pool)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~CHR, nrow = 5, ncol = 2, scales = "free_x") +
  theme_classic()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  geom_hline(yintercept = 0.5, linetype = "dashed", alpha = 0.5)

#ggsave("plots/SNP_density_plot_redo.jpg", plot = SNP_density_plot, height = 20, width = 12)

SNP_density_plot
```

```{r all on one axis fig.height=7, fig.width=25}
SNP_density_plot <- plot_all_SNP_density %>%
  ggplot(aes(x = bin, y = SNP_density, colour = pool)) +
  geom_point(alpha = 0.5) +
  facet_grid(~CHR, scales = "free_x", space = "free_x", switch = "x") +
  geom_hline(yintercept = 0.5, linetype = "dashed", alpha = 0.5) +
  theme_classic()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.spacing = unit(0.2, "lines"),
        legend.position = "bottom") +
  labs(x = "Chromosome",
       y = "SNP allele frequency")

#ggsave("plots/SNP_density_one_axis.jpg", SNP_density_plot, height = 7, width = 20)

SNP_density_plot
```


---

##  Plot Differences Between Pools

```{r fig.height=20, fig.width=12}
plot_snp_diff <- plot_all_SNP_density %>%
  pivot_wider(names_from = pool, values_from = SNP_density) %>%
  mutate(diff_snps = early - late) %>%
  pivot_longer(cols = c(early, late), names_to = "pool")

plot_snp_diff %>%
  ggplot(aes(x = bin, y = diff_snps)) +
  geom_point(alpha = 0.1) +
  geom_smooth() +
  facet_wrap(~CHR, nrow = 5, ncol = 2, scales = "free") +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5)

delta_SNP <- plot_snp_diff %>%
  dplyr::mutate(bin = as.numeric(levels(bin))[bin]) %>% 
  ggplot() +
  aes(x = bin, y = diff_snps) +
  geom_point(alpha = 0.1) +
  geom_smooth() +
  facet_wrap(~CHR, nrow = 5, ncol = 2, scales = "free_x") +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5)+
  scale_x_continuous(
    breaks = seq(0, 6e7, by = 5e6), # This is where you define your bin range
    labels = function(x) paste0(x / 1e6 , " Mb")  # Converts bins to Mb
  )

#ggsave("plots/delta_SNP_redo.jpg", plot = delta_SNP, height = 20, width = 12)
```

```{r all chr on one axis}
SNP_delta_one_axis <- plot_snp_diff %>%
  dplyr::mutate(bin = as.numeric(levels(bin))[bin]) %>% 
  ggplot() +
  aes(x = bin, y = diff_snps, colour = CHR) +
  geom_point(alpha = 0.5) +
  geom_smooth(colour = "black", alpha = 0.3, size = 0.4) +
  facet_grid(~CHR, scales = "free_x", space = "free_x", switch = "x") +
  geom_hline(yintercept = 0, linetype = "solid", alpha = 0.7)+
  geom_hline(yintercept = 0.45, linetype = "dashed", alpha = 0.5)+
  geom_hline(yintercept = -0.45, linetype = "dashed", alpha = 0.5)+
  theme_classic()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.margin = unit(0, "lines"),
        legend.position = "none") +
  scale_x_continuous(expand = c(0, 0)) +
  ylim(-0.6,0.6)+
  scale_y_continuous(breaks = seq(-0.6, 0.6, by = 0.3))+
  labs(x = "Chromosome",
       y = "SNP allele frequency difference (Δ-SNP)")
  
ggsave("plots/delta_SNP_one_axis.jpg", plot = SNP_delta_one_axis, height = 7, width = 20)

SNP_delta_one_axis
```


---


## Individual SNP delta-index

## Zoom In on a Chromosome and Identify Strong SNPs

```{r zoom in A01}
# Zoom in on A01

a01_dsnp <- plot_snp_diff %>%
  filter(CHR == "A01") %>%
  dplyr::mutate(bin = as.numeric(levels(bin))[bin]) %>% 
  ggplot() +
  aes(x = bin, y = diff_snps) +
  geom_point(alpha = 0.2) +
  geom_smooth() +
  geom_hline(yintercept = 0, linetype = "longdash", alpha = 0.5)+
  geom_hline(yintercept = 0.45, linetype = "dashed", alpha = 0.6, color = "red")+
  geom_hline(yintercept = -0.45, linetype = "dashed", alpha = 0.6, color = "red")+
  #annotate("rect", fill = "gray", alpha = 0.3, 
  #         xmin = -Inf, xmax = Inf,
  #         ymin = -0.45, ymax = 0.45)+
  scale_x_continuous(
    breaks = seq(0, 4e7, by = 5e6), # This is where you define your bin range
    labels = function(x) paste0(x / 1e6 , " Mb")  # Converts bins to Mb
  )+
  theme_classic()+
  labs(x = "Chromosome position",
       y = "Δ-SNP allele frequency")

a01_snp_diff <- plot_all_SNP_density %>%
  filter(CHR == "A01") %>% 
  dplyr::mutate(bin = as.numeric(levels(bin))[bin]) %>% 
  ggplot(aes(x = bin, y = SNP_density, colour = pool)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0.5, linetype = "longdash", alpha = 0.5) +
  theme_classic()+
  theme(legend.position = "right") +
  labs(x = "Chromosome position",
       y = "SNP allele frequency")+
    scale_x_continuous(
    breaks = seq(0, 4e7, by = 5e6), # This is where you define your bin range
    labels = function(x) paste0(x / 1e6 , " Mb")  # Converts bins to Mb
  )+
    scale_colour_manual(values=c("late"="#EE3377",
                               "early"="#0077BB"))

# Chromosome schematic as rectangle
snp_table <- read_tsv("output_tables/snp_table.tsv")

genes_a01 <- snp_table %>% 
  filter(CHR == "A01",
         IMPACT == "HIGH" | 
         ATH_NAME == "SUS2") %>% 
  select(POS, ATH_NAME)

highlight_genes <- c("SUS2","CKL6")

a01_chromo <- ggplot() +
  geom_rect(aes(xmin = 0, xmax = 4e7, ymin = 0, ymax = 0.5),
            fill = "white", color = "gray") +
  geom_segment(data = genes_a01,
               aes(x = POS, xend = POS, y = 0, yend = 0.5),
               color = "blue", linewidth = 0.3) +
    # Highlighted genes in red
  geom_segment(data = filter(genes_a01, ATH_NAME %in% highlight_genes),
               aes(x = POS, xend = POS, y = 0, yend = 0.5),
               color = "red",  linewidth = 0.5) +
  # Labels above for highlighted genes only
  geom_text(data = genes_a01 %>% filter(genes_a01$ATH_NAME %in% highlight_genes),
            aes(x = POS, y = -0.1, label = ATH_NAME),
            size = 3, vjust = 1,  color = "red", angle = 45, hjust = 1) +
  theme_minimal()+
  labs(x = "Chromosome A03", y = NULL)+
  scale_x_continuous(
    breaks = seq(0, 4e7, by = 5e6), # This is where you define your bin range
    labels = function(x) paste0(x / 1e6 , " Mb")) +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid = element_blank(),
    axis.title.y = element_blank(),
    axis.line.y = element_blank(),
    plot.title = element_blank()
  )+
    # Use simpler coordinate system
  coord_cartesian(ylim = c(-0.5, 0.6))

a01_plot <- a01_snp_diff / a01_dsnp / a01_chromo +
  plot_annotation(title = "ChrA01")

a01_plot

#ggsave("plots/a01_plot.jpg", plot = a01_plot, width = 10, height = 7)

#ggsave("plots/a01_snp_diff.jpg", plot = a01_snp_diff, width = 7, height = 5)
#ggsave("plots/a01_dsnp.jpg", plot = a01_dsnp, width = 7, height = 5)

snp_table %>% 
  filter(CHR == "A01",
         IMPACT == "HIGH" | 
         ATH_NAME == "SUS2") %>% 
  select(POS, ATH_NAME) %>% 
  dplyr::mutate(ATH_copy = NA) %>% 
  dplyr::mutate(ATH_copy = str_replace(as.character(POS), "6105582", "SUS2"))

```


```{r zoom in A02}
# Zoom in on A02

a02_dsnp <- plot_snp_diff %>%
  filter(CHR == "A02") %>%
  dplyr::mutate(bin = as.numeric(levels(bin))[bin]) %>% 
  ggplot() +
  aes(x = bin, y = diff_snps) +
  geom_point(alpha = 0.2) +
  geom_smooth() +
  geom_hline(yintercept = 0, linetype = "longdash", alpha = 0.5)+
  geom_hline(yintercept = 0.45, linetype = "dashed", alpha = 0.6, color = "red")+
  geom_hline(yintercept = -0.45, linetype = "dashed", alpha = 0.6, color = "red")+  scale_x_continuous(
    breaks = seq(0, 4e7, by = 5e6), # This is where you define your bin range
    labels = function(x) paste0(x / 1e6 , " Mb")  # Converts bins to Mb
  )+
  theme_classic()+
  labs(x = "Chromosome position",
       y = "SNP allele frequency difference (Δ-SNP)")

a02_snp_diff <- plot_all_SNP_density %>%
  filter(CHR == "A02") %>% 
  dplyr::mutate(bin = as.numeric(levels(bin))[bin]) %>% 
  ggplot(aes(x = bin, y = SNP_density, colour = pool)) +
  geom_point(alpha = 0.5) +
  #facet_grid(~CHR, scales = "free_x", space = "free_x", switch = "x") +
  geom_hline(yintercept = 0.5, linetype = "longdash", alpha = 0.5) +
  theme_classic()+
  theme(legend.position = "bottom") +
  labs(x = "Chromosome position",
       y = "SNP allele frequency")+
    scale_x_continuous(
    breaks = seq(0, 4e7, by = 5e6), # This is where you define your bin range
    labels = function(x) paste0(x / 1e6 , " Mb")  # Converts bins to Mb
  )+
      scale_colour_manual(values=c("late"="#EE3377",
                               "early"="#0077BB"))

a02_plot <- a02_snp_diff / a02_dsnp +
  plot_annotation(title = "ChrA02")

#ggsave("plots/a02_plot.jpg", plot = a02_plot, width = 10, height = 7)

#ggsave("plots/a02_snp_diff.jpg", plot = a02_snp_diff, width = 7, height = 5)
#ggsave("plots/a02_dsnp.jpg", plot = a02_dsnp, width = 7, height = 5)


# Identify SNPs in A01 with high SNP allele difference
p045 <- plot_snp_diff %>%
  filter(CHR == "A01") %>%
  select(bin, diff_snps) %>%
  distinct() %>%
  arrange(desc(diff_snps)) %>%
  filter(diff_snps > 0.45)

print(p045)


# Chromo A02
genes_a02 <- snp_table %>% 
  filter(CHR == "A02",
         IMPACT == "HIGH" | 
         ATH_NAME == "SUS2") %>% 
  select(POS, ATH_NAME)

highlight_genes <- c("SUS2","CKL6")

a01_chromo <- ggplot() +
  geom_rect(aes(xmin = 0, xmax = 4e7, ymin = 0, ymax = 0.5),
            fill = "white", color = "gray") +
  geom_segment(data = genes_a01,
               aes(x = POS, xend = POS, y = 0, yend = 0.5),
               color = "blue", linewidth = 0.3) +
    # Highlighted genes in red
  geom_segment(data = filter(genes_a01, ATH_NAME %in% highlight_genes),
               aes(x = POS, xend = POS, y = 0, yend = 0.5),
               color = "red",  linewidth = 0.5) +
  # Labels above for highlighted genes only
  geom_text(data = genes_a01 %>% filter(genes_a01$ATH_NAME %in% highlight_genes),
            aes(x = POS, y = -0.1, label = ATH_NAME),
            size = 3, vjust = 1,  color = "red", angle = 45, hjust = 1) +
  theme_minimal()+
  labs(x = "Chromosome A03", y = NULL)+
  scale_x_continuous(
    breaks = seq(0, 4e7, by = 5e6), # This is where you define your bin range
    labels = function(x) paste0(x / 1e6 , " Mb")) +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid = element_blank(),
    axis.title.y = element_blank(),
    axis.line.y = element_blank(),
    plot.title = element_blank()
  )+
    # Use simpler coordinate system
  coord_cartesian(ylim = c(-0.5, 0.6))
```


```{r Zoom in A03}
# Zoom in on A03

a03_dsnp <- plot_snp_diff %>%
  filter(CHR == "A03") %>%
  dplyr::mutate(bin = as.numeric(levels(bin))[bin]) %>% 
  ggplot() +
  aes(x = bin, y = diff_snps) +
  geom_point(alpha = 0.2) +
  geom_smooth() +
  geom_hline(yintercept = 0, linetype = "longdash", alpha = 0.5)+
  geom_hline(yintercept = 0.45, linetype = "dashed", alpha = 0.6, color = "red")+
  geom_hline(yintercept = -0.45, linetype = "dashed", alpha = 0.6, color = "red")+  scale_x_continuous(
    breaks = seq(0, 4e7, by = 5e6), # This is where you define your bin range
    labels = function(x) paste0(x / 1e6 , " Mb")  # Converts bins to Mb
  )+
  theme_classic()+
  labs(x = "Chromosome position",
       y = "SNP allele frequency difference (Δ-SNP)")

a03_snp_diff <- plot_all_SNP_density %>%
  filter(CHR == "A03") %>% 
  dplyr::mutate(bin = as.numeric(levels(bin))[bin]) %>% 
  ggplot(aes(x = bin, y = SNP_density, colour = pool)) +
  geom_point(alpha = 0.5) +
  #facet_grid(~CHR, scales = "free_x", space = "free_x", switch = "x") +
  geom_hline(yintercept = 0.5, linetype = "longdash", alpha = 0.5) +
  theme_classic()+
  theme(legend.position = "bottom") +
  labs(x = "Chromosome position",
       y = "SNP allele frequency")+
    scale_x_continuous(
    breaks = seq(0, 4e7, by = 5e6), # This is where you define your bin range
    labels = function(x) paste0(x / 1e6 , " Mb")  # Converts bins to Mb
  )+
      scale_colour_manual(values=c("late"="#EE3377",
                               "early"="#0077BB"))

a03_plot <- a03_snp_diff / a03_dsnp +
  plot_annotation(title = "ChrA03")

#ggsave("plots/a03_plot.jpg", plot = a03_plot, width = 10, height = 7)


#ggsave("plots/a03_snp_diff.jpg", plot = a03_snp_diff, width = 7, height = 5)
#ggsave("plots/a03_dsnp.jpg", plot = a03_dsnp, width = 7, height = 5)
```


```{r}
# Load libraries
library(ggplot2)
library(dplyr)
library(patchwork)  # For combining plots

# Simulated gene data
genes <- data.frame(
  START = c(5e6, 1.2e7, 1.9e7, 3.1e7),
  STOP = c(5.2e6, 1.25e7, 2e7, 3.2e7),
  LABEL = c("Gene1", "Gene2", "Gene3", "Gene4")
) %>%
  mutate(MID = (START + STOP)/2)


# Δ-SNP plot (your style)
a03_dsnp <- plot_snp_diff %>%
  filter(CHR == "A03") %>%
  ggplot(aes(x = bin, y = diff_snps)) +
  geom_point(alpha = 0.2) +
  geom_smooth() +
  geom_hline(yintercept = 0, linetype = "longdash", alpha = 0.5) +
  geom_hline(yintercept = 0.45, linetype = "dashed", alpha = 0.6, color = "red") +
  geom_hline(yintercept = -0.45, linetype = "dashed", alpha = 0.6, color = "red") +
  scale_x_continuous(
    limits = c(0, 4e7),
    breaks = seq(0, 4e7, by = 5e6),
    labels = function(x) paste0(x / 1e6, " Mb")
  ) +
  theme_classic() +
  labs(x = "Chromosome position", y = "Δ-SNP index")

# Chromosome schematic as rectangle
a03_chromo <- ggplot() +
  geom_rect(aes(xmin = 0, xmax = 4e7, ymin = 0, ymax = 0.5),
            fill = "white", color = "gray") +
  geom_segment(data = genes,
               aes(x = MID, xend = MID, y = 0, yend = 0.5),
               color = "blue", size = 1) +
  geom_text(data = genes,
            aes(x = MID, y = 0.6, label = LABEL),
            size = 3, vjust = 0) +
  # Use simpler coordinate system
  coord_cartesian(ylim = c(0, 1)) +
  theme_void() +
  labs(title = "Chromosome A03 with Gene Locations") +
  theme(plot.title = element_text(hjust = 0.5))


# Combine vertically with patchwork
combined_plot <- a03_chromo / a03_dsnp + plot_layout(heights = c(1, 3))

# Show plot
print(combined_plot)

```


## 12. Sarisha SNP distribution

```{r sarisha SNP}
sarisha_filtered <- Sarisha_snps_only %>%
  rename(CHR = `#CHROM`) %>%
  mutate(CHR_POS = paste(CHR, POS, sep = "_")) %>%
  semi_join(good_snps, by = "CHR_POS")

calc_allele_freq(sarisha_filtered)


```
```{r}
# 1. Read in alternate-parent SNPs
#alt_snps <- read_tsv("alternate_parent_snps.tsv", col_names = c("Chr", "Pos"))

alt_snps <- good_snps %>% 
  separate(col = CHR_POS,
           into = c("Chr","Pos"),
           sep = "_",
           remove = T) %>% 
  dplyr::mutate(Pos = as.numeric(Pos)) %>% 
  filter(Chr %in% rapa_chr)
  

# 2. Bin the genome into windows
bin_size <- 1e6  # 1 Mb bins

snp_bins <- alt_snps %>%
  mutate(Bin = (Pos %/% bin_size) * bin_size) %>%  # Integer division to assign to bins
  group_by(Chr, Bin) %>%
  summarize(SNP_Count = n(), .groups = "drop")

# 3. Plot the SNP density
snp_density_bar <- ggplot(snp_bins, aes(x = Bin / 1e6, y = SNP_Count, fill = SNP_Count == 0)) +
  geom_col(width = 1) +
  scale_fill_manual(values = c("steelblue", "red"), guide = "none") +
  facet_wrap(~ Chr, scales = "free_x", space = "free_x") +
  labs(x = element_blank(), y = "SNP count",
       title = "SNP Density of Sarisha SNPs") +
  theme_classic() +
      theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.margin = unit(0.2, "lines")
      )

snp_density_heat <- ggplot(snp_bins, aes(x = Bin / 1e6, y = 0, fill = SNP_Count)) +
  geom_tile(height = 0.8, width = 0.8) +
  scale_fill_viridis_c(option = "plasma", trans = "sqrt") +
  facet_wrap(~ Chr, scales = "free_x", switch = "x") +
  scale_x_continuous(
    breaks = seq(0, max(snp_bins$Bin / 1e6), by = 5))+
  labs(x = "Chromosome", y = element_blank()) +
  theme_classic() +
      theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.margin = unit(0.2, "lines"),
        legend.position = "none"
      )

snp_density_plot <- snp_density_bar /
  snp_density_heat

```

```{r}
snp_bins %>% 
  summary()
```

```{r combine plots}
snp_density_barheat <- ggplot(snp_bins, aes(x = Bin / 1e6, y = SNP_Count, fill = SNP_Count)) +
  geom_col(width = 1.2) +
  scale_fill_viridis_c(option = "plasma", trans = "sqrt") +
  scale_y_continuous(trans = "sqrt") +
  scale_y_continuous(
  trans = "sqrt",
  breaks = c(1000, 5000))+
  facet_grid(~ Chr, scales = "free_x", space = "free_x") +
  geom_hline(yintercept = 512, linetype = "dashed", alpha = 0.3)+
  labs(x = element_blank(), y = "SNP count",
       title = element_blank()) +
  theme_classic() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )

SNP_density_plot <- plot_all_SNP_density %>%
  ggplot(aes(x = bin, y = SNP_density, colour = pool)) +
  geom_point(alpha = 0.5) +
  facet_grid(~CHR, scales = "free_x", space = "free_x", switch = "x") +
  geom_hline(yintercept = 0.5, linetype = "longdash", alpha = 0.5) +
  theme_classic()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.spacing = unit(0.2, "lines"),
        strip.background = element_blank(),
        strip.text.x = element_blank()) +
  labs(x = element_blank(),
       y = "SNP allele frequency")

SNP_delta_one_axis <- plot_snp_diff %>%
  dplyr::mutate(bin = as.numeric(levels(bin))[bin]) %>% 
  ggplot() +
  aes(x = bin, y = diff_snps, colour = CHR) +
  geom_point(alpha = 0.5) +
  geom_smooth(colour = "black", alpha = 0.3, size = 0.4) +
  facet_grid(~CHR, scales = "free_x", space = "free_x", switch = "x") +
  geom_hline(yintercept = 0, linetype = "longdash", alpha = 0.5)+
  geom_hline(yintercept = 0.45, linetype = "dashed", alpha = 0.6, color = "red")+
  geom_hline(yintercept = -0.45, linetype = "dashed", alpha = 0.6, color = "red")+
  theme_classic()+
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.spacing = unit(0.2, "lines"),
        legend.position = "none"
        ) +
  scale_x_continuous(expand = c(0, 0)) +
  ylim(-0.6,0.6)+
  scale_y_continuous(breaks = seq(-0.6, 0.6, by = 0.3))+
  labs(x = "Chromosome",
       y = "Δ-SNP allele frequency")
  

combine_plot <- snp_density_barheat /
  SNP_density_plot +
  SNP_delta_one_axis +
  plot_layout(widths = c(1, 1, 1), heights = unit(c(1, 5, 5), c('null', 'null', 'null')))+
  plot_annotation(tag_levels = 'A')
  
ggsave("plots/sarisha_snp_density_plot.jpg", combine_plot, height = 8, width = 17)
```

